Modeling Interventions

### 2021-03-19 paper/paper.Rmd
more tweaks

Accumulator models have already contributed significantly to several theoretical issues. 
For example, @mcmurray2007 elegantly demonstrated that children’s "vocabulary explosion" -- an acceleration in vocabulary growth in the second year -- can result from the steady accumulation of word tokens without changes in the environment or learning mechanism. 
Other work has examined similar models using more realistic distributions of words (i.e., Zipf-distributed), developmental change in learning mechanisms, and quantitative comparison to children’s aggregate vocabulary growth [@mayor2010; @hidaka2013; @mollica2017]. 
Although this work is exciting, these models have not yet made direct predictions about learning in individual children [@bergelson2020comprehension].

One difficulty in connecting such models to data is that the relevant variables are often measured in relative, rather than absolute, units. 
This practice is common. 
For example, measures of intelligence are given on a standardized scale defined by population variability. 
Yet in early language -- unusually in a psychological domain -- we have access to absolute units.
We can count how many words a child hears and express this number as a rate [e.g., words per hour; @cristia2020accuracy]. 
We can similarly estimate how many words they know [e.g., by parents' reports of vocabulary size; @frank2021]. 
These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations [@dailey2021language]. 
Few datasets have measurements of the relevant variables, however. 
Thus, one upshot of our argument here is that future data collection efforts should attempt whenever possible to report measurements in absolute units.

@@

Since seminal work by @hart1995, the connection between children’s language input and the growth of vocabulary has been a topic of intense interest. 
Numerous studies have reported positive associations between these variables [e.g., @hoff2003specificity; @weisleder2013] -- as predicted by accumulator models -- though their magnitude varies across studies [@wang2020meta].
Such correlations are also partially moderated by other factors, including socioeconomic [@hoff2003specificity; @dailey2021language] and genetic [@hayiou2014language] variables.
Randomized interventions are the gold standard for measuring causal effects of parental language input on language learning, although such studies are costly and difficult to conduct. 
When such studies are conducted, they show modest but reliable effects [e.g., @suskind2016project], providing support for a causal connection between language input and outcomes. 
Critically, however, these studies only estimate the effect of the change in input on variation in outcome, hence, providing only a relative, rather than an absolute, estimate of how important input is to vocabulary learning.

@@

To answer this question, we fit IRT models to parent report data about the words that children produce. 
We downloaded data from 5,429 monolingual, English-learning children between 16 and 30 months from Wordbank [@frank2021]. 
These data were collected using the MacArthur-Bates Communicative Development Inventory (CDI), a reliable and valid parent report instrument; they include 680 binary responses per child about whether a particular child produced a particular word at a given age. 
Each model estimated parameters for each child's linguistic ability and each word's difficulty.

@@

```{r, out.width="1.0\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Predicted acquisition curves from a fitted English IRT model as a function of lexical class and the number of expected tokens per month (left), and for a sample of the 680 words on the CDI (right)."), echo=FALSE}
knitr::include_graphics("./figs/fig3_predicted_age_LC_and_items.pdf")
```

@@

**Leveraged learning and the role of processing**. 
As our results above show, older children accumulate language from their input faster than younger children.
Why?
Are older children simply better at remembering words, leading to "bigger drops" in the bucket for older children? 
Or children *leverage* their knowledge of language to learn faster [@mitchell2009]? 
They could do this by reasoning about new words by exclusion [@markman1988children]. 
Or their increasing fluency with the words they know could help them learn new words. A surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which they process familiar words [e.g., @fernald2006picking; @marchman2016early].
These theories predictions about which words should faciliate which others predictions that could be tested using the models we describe.

@@

**Beyond vocabulary**. 
We have described a model of the accumulation of words.
Yet language is a rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings. 
In early views, syntax and morphology were conceptualized as distinct and unconnected, but this conception has not been borne out empirically. 
Instead, evidence shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@bornstein2013language; @frank2021]. 
Theoretically, accumulator models like the standard model are generic models of skill acquisition. 
If language learning is a form of skill acquisition [@chater2018language], such connections could lead the way towards extensions of the standard model to the accumulation of broader units of language-like constructions.

>>>>

Accumulator models have already contributed significantly to several theoretical issues. 
For example, @mcmurray2007 elegantly demonstrated that children’s "vocabulary explosion" -- an acceleration in vocabulary growth in the second year -- can result from the steady accumulation of word tokens without changes in the environment or learning mechanism. 
Other work has examined similar models using more realistic distributions of words, developmental change in learning mechanisms, and comparison to children’s aggregate vocabulary growth [@mitchell2009; @hidaka2013; @mollica2017]. 
Although this work is exciting, these models have not yet made direct predictions about learning in individual children [@bergelson2020comprehension].

One difficulty in connecting such models to data is that the relevant variables are often measured in relative, rather than absolute, units. 
This practice is common. 
For example, measures of intelligence are given on a standardized scale defined by population variability. 
Yet in early language -- unusually in a psychological domain -- we have access to absolute units.
We can count how many words a child hears and express this number as a rate [e.g., words per hour; @cristia2020accuracy]. 
We can similarly estimate how many words they know [e.g., by parents' reports of vocabulary size; @frank2021]. 
These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations. 
Few datasets have measurements of the relevant variables, however. 
Thus, one upshot of our argument here is that future data collection efforts should attempt whenever possible to report measurements in absolute units.

@@

Since seminal work by @hart1995, the connection between children’s language input and the growth of vocabulary has been a topic of intense interest. 
Numerous studies have reported positive associations between these variables [e.g., @hoff2003specificity] -- as predicted by accumulator models -- though their magnitude varies across studies [@wang2020meta].
Such correlations are also partially moderated by other factors, including socioeconomic [@hoff2003specificity] and genetic [@hayiou2014language] variables.
Randomized interventions are the gold standard for measuring causal effects of parental language input on language learning, although such studies are costly and difficult to conduct. 
When such studies are conducted, they show modest but reliable effects [e.g., @suskind2016project], providing support for a causal connection between language input and outcomes. 
Critically, however, these studies only estimate the effect of the change in input on variation in outcome, hence, providing only a relative, rather than an absolute, estimate of how important input is to vocabulary learning.

@@

To answer this question, we fit IRT models to parent report data about the words that children produce.
In the simple setting we describe, these models can be written as multi-level logistic regression models [@de-boeck2011], albeit with a clearer theoretical interpretation than typical regression models. 
We downloaded data from 5,429 monolingual, English-learning children between 16 and 30 months from Wordbank [@frank2021]. 
These data were collected using the MacArthur-Bates Communicative Development Inventory (CDI), a reliable and valid parent report instrument; they include 680 binary responses per child about whether a particular child produced a particular word at a given age. 
Each model estimated parameters for each child's linguistic ability and each word's difficulty.

@@

```{r, out.width="1.0\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Predicted acquisition curves derived from a fitted age-dependent accumulator model (see text) as a function of lexical class and the number of expected tokens per month (left), and for a sample of specific (right)."), echo=FALSE}
knitr::include_graphics("./figs/fig3_predicted_age_LC_and_items.pdf")
```

@@

**Leveraged learning and the role of processing**. 
As our results above show, older children accumulate language from their input faster than younger children.
Why?
Are older children simply better at remembering words, leading to "bigger drops" in the bucket for older children? 
Or children *leverage* their knowledge of language to learn faster [@mitchell2009]? 
They could do this by reasoning about new words by exclusion [@markman1988children]. 
Or their increasing fluency with the words they know could help them learn new words. A surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which they process familiar words [e.g., @marchman2016early].
These theories predictions about which words should faciliate which others predictions that could be tested using the models we describe.

@@

**Beyond vocabulary**. 
We have described a model of the accumulation of words.
Yet language is a rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings. 
In early views, syntax and morphology were conceptualized as distinct and unconnected, but this conception has not been borne out empirically. 
Instead, evidence shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@frank2021]. 
Theoretically, accumulator models like the standard model are generic models of skill acquisition. 
If language learning is a form of skill acquisition [@chater2018language], such connections could lead the way towards extensions of the standard model to the accumulation of broader units of language-like constructions.
